\input{Figures/figure3}
Murugan, Huse, and Leibler recently discovered that energetic discrimination in a general network requires a {\it discriminatory fence}~\cite{Murugan2014}, which can be idealized as a ladder graph having two sides, each with $N$ loops (Figure \ref{fig:ladder}(a)).  The sides of the ladder are symmetric about the $0$ node; the network aims to discriminate between states represented by its upper corners (i.e., $x_{s2}$ in Figure \ref{fig:ladder}(a)).  Rate constants $u^S,\ d^S, \ S=\{W, R\}$ will differ for the `Wrong' ($W$) and `Right' ($R$) sides of the network.

The ladder idealization captures the fact that a general energetic discrimination network must have dominant `forward' ($f$) and `reverse' ($b$) paths which are parallel to each other and effectively one-directional.  On the pathway towards the product state, there is the constant threat of `discard' ($d$), after which the reaction is exposed to a one-directional pathway away from the product state ($b$).  There is also the possibility of `rescue' ($u$) from discard. 

The Kramer's form rate constants for this network are
\begin{center}
\begin{tabular}{ll}
$u^R = \omega_d e^{\epsilon_u+\delta}$ & $d^R = \omega_d e^{\delta}$\\
\\
$u^W =  \omega_d e^{\epsilon_u} $&$ d^W = \omega_d e^{\gamma}$.\\
\end{tabular}
\end{center}
And there is no discrimination ($f^R=f^W=f$) along the forward or reverse pathways:
\[
\begin{aligned}
f = \omega_f & & b = \omega_b,\\
\end{aligned}
\]
which we approximate to be one-directional for analytical convenience, but treat as bidirectional with small reverse rates when necessary for computing dissipation.

It is clear from the Kramer's form constants that to discriminate in the energetic regime (i.e., via $\gamma$), a high discard rate ($d$) is required.  Indeed, the error rate for an $N$-loop network~\footnote{An $N$-loop network will strictly speaking be composed of $2N+1$ loops, $N$ on each side of the ladder and a single reactant node.} in this regime is 
\begin{equation}\label{energy_error}
\xi_{\text{energetic}} = \frac{1}{e^\gamma}\left (\frac{\omega_d+\omega_f}{\omega_d e^{\gamma}+\omega_f} \right )^N
\end{equation}
which achieves its minimum when discards are high relative to steps toward reaction completion:
\begin{equation}\label{elim}
\omega_d/\omega_f \to \infty.
\end{equation}
Discrimination in this regime is fundamentally processive, and global: accuracy relies on sequential exposure to frequently realized discard pathways, and {\it each} reaction step contributes to discrimination via the potential for discard.  Correspondingly, orthogonality monotonically decreases in the Equation \ref{elim} limit (Appendix \ref{app:ladder_error}), and is minimized in the high discard regime (Figure \ref{fig:ladder}(b), solid lines).

In contrast, we find that the kinetic regime has error fraction given by  (Appendix~\ref{app:ladder_error}):
\begin{equation}\label{general:kinetic_error}
\xi_{\rm kinetic} =  \frac{(\phi+1)^\alpha(1+\eta e^\delta)^\alpha}{(\phi e^\delta+1)^\alpha(\eta+1)^\alpha}.
\end{equation}
where 
\[
\phi = \omega_d e^{\epsilon_u}/\omega_b, \ \ \ and \ \ \ \eta =\omega_d/ \omega_f.
\]
The error $\xi_{\text{kinetic}}$ is minimized when $\eta\to0$ and $\phi\to\infty$, which is to say that:
\begin{equation}\label{kin_lad}
\omega_d/\omega_f \to 0, \ \ \omega_de^{\epsilon_u}/\omega_b \to \infty.
\end{equation}
These limits imply that network dynamics are being pushed quickly towards the final product nodes ($\omega_f,\ \epsilon_u$ large, $\omega_b$ small).  This makes local discrimination possible; and indeed orthogonality is monotonically increasing in the Equation \ref{kin_lad} limit (Appendix \ref{app:ladder_orth}).

Quick movement towards final product nodes is in opposition to high discard rates; we can thus summarize the difference between the energetic and kinetic regimes by observing their difference with respect to the discard rate ($d$, Figure \ref{fig:ladder}(b) x-axis), which reveal the expected orthogonality-error relationships in the two regimes.  Note that these limits correspond to the dynamical phase localization limits described in \cite{Murugan2016}.

We are now in a position to understand the orthogonality of this model in terms of its effective pathways towards the final product nodes.  The energetic discrimination requirement that $f<<d$ means that the network effectively contains only a single pathway to the product.  Intuitively, the single pathway results from the slowness of one-directional progress towards the final product; rescue pathways cannot add additional paths to the final product because they are effectively equilibrated relative to the slow forward progress.  Corresponding to this intuition, we find analytically that $u, \ b,$ have essentially no effect on orthogonality in the $f<<d$ regime (Appendix \ref{app:ladder_orth}).  This argument is consistent with the fact that the discrimination error in the energetic regime (Equation \ref{energy_error}) is independent of $u, \ b,$ but in the kinetic regime, which requires $d<<f,$ we find that $u, b$ are important factors in the error expression (Equation \ref{general:kinetic_error}) and orthogonality requirements (Equation \ref{kin_lad}).

In the energetic regime, we observe that as $f$ becomes close to $d$ (red tick, Figure \ref{fig:ladder}(c)), orthogonality rises sharply Figure \ref{fig:ladder}(c).  We understand this to result from many more effective pathways now leading to the final product.  Again, the rise in orthogonality as we increase $f$ leads to the non-monotonic behavior of the error rate.  

Our understanding of orthogonality in terms of effective pathways allows us to apply thermodynamic drive in the energetic regime such that drive does {\it not} increase orthogonality.  Our arguments above state that $f << d,$ enforces the single pathway and hence orthogonal regime.  Therefore, if we dissipate energy to drive $d,$ we should find that the orthogonality decreases, and indeed we do (Figure \ref{fig:ladder}(d)).  Note that Figure \ref{fig:ladder}(c) was generated with the same parameters as Figure \ref{fig:ladder}(d); all that's changed is the reaction we choose to drive.  In this parametric limit, the orthogonality and dissipation requirements are not contravening.  

\input{Figures/figure4}
Finally, we note that (as in the Hopfield-Ninio regime) highly selective  - seven orders of magnitude - dissipation driven product switching is possible between states which are favored by different energy types (Figure \ref{fig:ladder_switch}).


