We first demonstrate the relationship between orthogonality and discrimination using the classical Hopfield-Ninio scheme, shown graphically in Figure \ref{fig:hopfield}(a).  Here, substrates $S= \{W, \ R\}$ compete to form complexes with enzyme $E$.  `Wrong' and `Right' products are formed from substrates $R$ and $W$ (respectively), at rates proportional to the steady state occupancy of the final complex $\rho_{ES}.$  We thus define the {\it error fraction} achieved by the discrimination scheme to be
\[
\xi = \frac{\rho_{EW}}{\rho_{ER}}.
\]
\input{Figures/figure1}
Ninio and Hopfield designed this scheme to amplify differences in the binding energies of $EW$ and $ER$ complex formation.  It is instructive to write the rate constants in Kramer's form.

We have for the $EW$ reactions:
\begin{center}
\begin{tabular}{ll}
$k^{\prime}_W = \omega e^\epsilon$ & $l^{\prime}_W = \omega_p$\\
\\
$k_W =  \omega e^\gamma $&$ l_W = \omega_p e^{\epsilon_p+\gamma}$\\
\end{tabular}
\end{center}
where: $\omega, \ \omega_p$ set overall rates; $\epsilon, \epsilon_p$ represent the enthalpy differences between free enzyme and complexes $ES^*, \ ES;$ and $\gamma$ represents the binding energy difference between right and wrong complexes.

The $ER$ reactions are given by:
\begin{center}
\begin{tabular}{ll}
$k^{\prime}_R = \omega e^{\epsilon+\delta}$ & $l^{\prime}_R = \omega_pe^{-\delta_p}$\\
\\
$k_R =  \omega e^\delta $&$ l_R = \omega_p e^{\epsilon_p-\delta_p}$\\
\end{tabular}
\end{center}
where $\delta$ sets the activation energy differences between right and wrong complexes.

There is no discrimination along the transitions between the intermediary and final product:
\begin{center}
\begin{tabular}{ccc}
$m = \omega_i,$  & & $m' = \omega_i e^{\eps_i}$.\\
\\
\end{tabular}
\end{center}

We begin by considering the relationship between error and orthogonality in the regime which is governed only by binding energy differences ($\gamma >0, \ \delta=0$), termed the `energetic regime.'  The Hopfield-Ninio scheme was originally designed for discrimination in this regime: the intermediary complex ($ES^*$) and dissipative drive ($m'$) introduce a delay which amplifies the fact that $EW$ complex formation has a faster off-rate (by factor of $\gamma$) relative to $ER$ formation.  

Simulations reveal that low orthogonality is necessary, but not sufficient, for low error rates in the energetic regime (Figure  \ref{fig:hopfield}(b)).  Analytically, we recall the long appreciated fact that the (energetic) Hopfield-Ninio scheme requires the parametric limit
\[
\frac{\omega_p}{\omega e^{\epsilon}}\rightarrow0 
\]
in order to minimize its error, and demonstrate that orthogonality is monotonically decreasing as this limit is approached (Appendix~\ref{app:hopfield}).

A less well-appreciated requirement for energetic discrimination concerns the nonequilibrium drive, represented by $m'.$  Some amount of drive is crucial for the discrimination scheme to work at all, but too much drive destroys discrimination \cite{Wong2018}.  We can understand this nonlinearity in terms of orthogonality (Figure  \ref{fig:hopfield}(c)).  Energy dissipation~\footnote{Throughout this text dissipation is calculated as the entorpy production rate, \unexpanded{$\Delta S_i = \frac{1}{2}\sum_{i,j}(k_{ij}\rho_j-k_{ji}\rho_i)\ln{\frac{k_{ij}\rho_j}{k_{ji}\rho_i}}$}} is helpful for discrimination up {\it until} the point at which it begins to drive up orthogonality.

We next turn to the regime governed by only activation energy differences ($\gamma =0, \ \delta > 0$), termed the `kinetic regime.'  Simulations reveal a bound opposite to that of the energetic regime: high orthogonality is necessary (but not sufficient) for low error (Supplemental Figure \ref{sfig:kin_sim}).  Analytically, we can derive the error in this regime to be
\[
\xi_{\text{kinetic}} = \frac{1+e^{-\delta}\eta b+e^{-2\delta} \eta c}{1+\eta b+\eta c}
\]
where 
\[
a = \omega\omega_i, \ \ \ b = \omega\omega_p, \ \ \ c = \omega_p\omega_i e^{\epsilon_i}, \ \ \ \eta=e^{\epsilon_p}/a.
\]
The $\xi_{\text{kinetic}}$ is minimized when $\eta\gg1$ and $c\gg b.$  That is, when there exists high drive ($\omega_i e^{\epsilon_i}\gg\omega)$ and free enthalpy product differences ($\epsilon_p\gg0$).  We demonstrate that orthogonality is monotonically {\it increasing} as these limits are approached (Appendix~\ref{app:hopfield}).

Differences between the two discriminatory  regimes are summarized in Figure  \ref{fig:hopfield}(d).  Increasing the dissipative drive ($\eps_i$) increases orthogonality, which allows for kinetic discrimination but precludes energetic discrimination.
\input{Figures/figure2}
The ability to modulate orthogonality via driving the single reaction $m'$ suggests a simple strategy for dissipation-driven product switching.
If products $EW, \ ER$ are favored by different energy types, they can be selected between only by driving $m'$ such that the network moves from low to high orthogonality. We achieve a four order of magnitude selection effect via this scheme (Figure~\ref{fig:hop_switch}).

Because the Hopfield-Ninio scheme only has one intermediary product, it is difficult to interpret in terms of the number of effective pathways towards the discriminatory products.  In order to illustrate the connection between discrimination, effective pathways, and orthogonality, we turn to a more general setting.